---
title: "Titanic.Rmd"
output: html_document
---

```{r}
setwd('/Users/evdokiapopova/Documents/data science club/kaggle_titanic')
library(ggplot2)
trainData <- read.csv('train.csv')
testData <- read.csv('test.csv')

summary(trainData)
summary(testData)
```
```{r}
names(trainData)
```

Let's start plotting the features

```{r}

qplot(x = Sex, data = trainData,color = I('black'), fill = I('blue')) + facet_wrap(~Survived)

```
More female survived than male.

```{r}

qplot(x = Age, data = trainData,color = I('black'), fill = I('blue')) + facet_wrap(~Survived)

```
Twice as more people of ages 20-30 didn't survive.

```{r}
qplot(x = Age, data = trainData,color = I('black'), fill = I('blue')) + facet_grid(Sex~Survived)
```

A lot of males of ages 20-40 didn't survive.


```{r}

qplot(x = SibSp, data = trainData,color = I('black'), fill = I('blue')) + facet_wrap(~Survived)

```
Almost twice as many "single" people died compared to survived. Let's take a look who they were.

```{r}
qplot(x = Age, data = trainData,color = I('black'), fill = I('blue')) + facet_grid(SibSp~Survived)

# lets take a look only at the people traveling alone

subsetSibSp_0 = subset(trainData, SibSp == 0)

qplot(x = Age, data = subsetSibSp_0,color = I('black'), fill = I('blue')) + facet_grid(Sex~Survived)

```


```{r}
qplot(x = Pclass, data = trainData,color = I('black'), fill = I('blue')) + facet_wrap(~Survived)

```
```{r}
qplot(x = Embarked, data = trainData,color = I('black'), fill = I('blue')) + facet_wrap(~Survived)

```
```{r}
qplot(x = Fare, data = trainData,color = I('black'), fill = I('green')) + facet_wrap(~Survived) + 
  scale_x_continuous(limits = c(0,300))


```

Combine both training and test data to clean up

```{r}
# merge both datasets but before create Survived column in test data with NA values
testData$Survived <- NA
allData <- rbind(trainData, testData)

head(allData)
tail(allData)
```
```{r}
# check missing values
sapply(allData,function(x) sum(is.na(x))) # lots of ages missing and 1 Fare
sapply(allData, function(x) sum(x=='')) # 2 embarked values are empty 

# unique values of each feature 
sapply(allData, function(x) length(unique(x)))

```

```{r}
# lets fill it with some values

# filling age and fare with average values

allData$Age[is.na(allData$Age)] <- round(mean(allData$Age,na.rm=T))
allData$Fare[is.na(allData$Fare)] <- round(mean(allData$Fare,na.rm=T))

# lots of people Embarked at S, so we fill those with S

sapply(allData, class)

allData$Embarked[allData$Embarked == ''] <- as.factor('S')

```

```{r}

#allData$Gender <- ifelse(allData$Sex == 'male', 1, 0)

allData = subset(allData, select = -c(1,4,9,11))

head(allData)
```

```{r}
# split back data

trainData <- allData[1:891,]
testData <- allData[892:nrow(allData),]
```


```{r}
library(Hmisc)

#cor(trainData$Age, trainData$Survived)

```


```{r}
library(corrgram)
corrgram(trainData, order=NULL, lower.panel=panel.shade,
  upper.panel=NULL, text.panel=panel.txt,
  main="correlation map of titanic data")
```
```{r}
library(corrplot)

#cor_data <- cor(trainData)
#head(trainData)

#corrplot(cor_data, method="color", use.names = TRUE)

#names(trainData)
```


Now let's try to build a model using Logistic regression.

**************************************
Logistic regression
**************************************

```{r}
model <- glm(Survived ~ Pclass + Sex + Age + Fare +Embarked ,family=binomial(link='logit'),data=trainData)

summary(model)
```

Let's test our model

```{r}
model_predict <- predict(model,newdata=subset(testData,select=c(Pclass, Sex, Age, Fare, Embarked),type='response'))  

#0.78469 score

summary(model_predict)

model_predict <- ifelse(model_predict>=0.5,1,0)
head(model_predict)

```
```{r}

predict_submit <- subset(testData, select = c(1))
head(predict_submit)
predict_submit$Survived <- model_predict

write.csv(predict_submit, file = "LogitPredict2.csv")
```

***************************************
RANDOM FOREST
***************************************

```{r}
library(party)
library(randomForest)
set.seed(415)

fit <- randomForest(Survived ~ Pclass + Age + SibSp + Parch + Fare +
                                            Embarked + Sex,
                      data=trainData, method = "class",
                      importance=TRUE, 
                      ntree=2000)

varImpPlot(fit)

Prediction <- predict(fit, testData)

Prediction <- ifelse(Prediction>=0.5,1,0)


## submission file

predict_submit$Survived <- Prediction

write.csv(predict_submit, file = "RandomForest1.csv") ## 0.77033 score

```


```{r}
### new random forest

fit <- cforest(as.factor(Survived) ~ Pclass + Age + SibSp + Parch + Fare + Embarked + Sex,data = trainData, controls=cforest_unbiased(ntree=2000, mtry=3))

Prediction <- predict(fit, testData, OOB=TRUE, type = "response")

## submission file

predict_submit$Survived <- Prediction

write.csv(predict_submit, file = "RandomForest2.csv") ## 0.77990 score

```


